QRF Covariate Selection

### 2020-02-04 analysis/R_scripts/fit_qrf_model.Rmd
revised the fit_qrf_model script, including choosing a different subset of metrics. Re-saved all the .rds objects.

All sites fell within the range of steelhead, but many of them were outside the range of spring/summer Chinook salmon. However, defining the range of sp/su Chinook is tricky, and it is unclear if the reported data for Chinook salmon includes only sites considered with the range of Chinook, or if we should exclude data from some sites based on our best understanding of the species' range. Currently, we are excluding sites outside the Chinook domain. We created our own list of sites in the John Day that were within a Chinook domain, based on which sites Chinook were found at some point (and downstream of those).

@@

# which of those sites are in Chinook domain?
chnk_sites = chnk_samps %>%
  as_Spatial() %>%
  snapPointsToLines_v2(points = .,
  # maptools::snapPointsToLines(points = .,
                              lines = chnk_domain %>%
                                mutate(id = 1:n()) %>%
                                select(id, MPG) %>%
                                as_Spatial(),
                              maxDist = max_snap_dist,
                              withAttrs = TRUE,
                              idField = 'id') %>%
  as('sf') %>%
  select(-nearest_line_id, -snap_dist) %>%
  # include some sites in the John Day where Chinook were found (or seemed to be close to sites where Chinook were found)
  rbind(st_read('../data/derived_data/Chnk_JohnDay_TrueObs.shp',
                quiet = T) %>%
          st_transform(st_crs(chnk_domain)) %>%
          select(-in_range))

fish_hab %<>%
  filter(Site %in% chnk_sites$Site)
  # filter(Species == 'Steelhead' |
  #          (Species == 'Chinook' & Site %in% chnk_sites$Site))

@@

```{r habitat-covariates}
sel_hab_mets = tibble(Species = c('Chinook'),
                      Metric = c('UcutArea_Pct',
                                 'FishCovNone',
                                 'SubEstGrvl',
                                 'FstTurb_Freq',
                                 'FstNT_Freq',
                                 'CU_Freq',
                                 'SlowWater_Pct',
                                 'NatPrin1',
                                 'DistPrin1',
                                 'avg_aug_temp',
                                 # 'Sin_CL',
                                 'Sin',
                                 'WetWdth_CV',
                                 'WetBraid',
                                 'WetSC_Pct',
                                 'Q',
                                 'WetWdth_Int',
                                 'LWFreq_Wet',
                                 'LWVol_WetFstTurb'))

@@

pred_hab_sites_chnk = pred_hab_sites %>%
  filter(!is.na(LON_DD)) %>%
  st_as_sf(coords = c('LON_DD', 'LAT_DD'),
           crs = 4326) %>%
  st_transform(crs = st_crs(chnk_domain)) %>%
  as_Spatial() %>%
  snapPointsToLines_v2(chnk_domain %>%
  # maptools::snapPointsToLines(chnk_domain %>%
                                mutate(id = 1:n()) %>%
                                select(id, MPG) %>%
                                as_Spatial(),
                              maxDist = snap_dist,
                              withAttrs = T,
                              idField = 'id') %>%
  as('sf') %>%
  as_tibble() %>%
  select(-nearest_line_id, -snap_dist, -geometry) %>%
  # add sites in the John Day
  bind_rows(st_read('../data/derived_data/Chnk_JohnDay_TrueObs.shp',
                    quiet = T) %>%
              as_tibble() %>%
              select(Site) %>%
              inner_join(pred_hab_sites))

@@

# compare range of covariates from model dataset and prediction dataset
range_comp = bind_rows(gaa_all %>%
                         filter(!Site %in% unique(pred_hab_sites$Site)) %>%
                         select(Site,
                                one_of(gaa_num)) %>%
                         gather(Metric, value, -Site) %>%
                         mutate(Source = 'non-DASH Sites'),
                       gaa_all %>%
                         filter(Site %in% unique(pred_hab_sites$Site)) %>%
                         select(Site,
                                one_of(gaa_num)) %>%
                         distinct() %>%
                         gather(Metric, value, -Site) %>%
                         mutate(Source = 'DASH Sites')) %>%
  mutate_at(vars(Source, Metric),
            list(as.factor))

range_max = range_comp %>%
  group_by(Metric, Source) %>%
  summarise_at(vars(value),
               tibble::lst(min, max),
               na.rm = T) %>%
  filter(Source == 'DASH Sites') %>%
  ungroup() %>%
  gather(type, value, -Metric, -Source)

covar_range_p = range_comp %>%
  ggplot(aes(x = Source,
             y = value,
             fill = Source)) +
  geom_boxplot() +
  facet_wrap(~ Metric,
             scales = 'free') +
  geom_hline(data = range_comp %>%
               group_by(Metric, Source) %>%
               summarise_at(vars(value),
                            tibble::lst(min, max),
                            na.rm = T) %>%
               filter(Source == 'DASH Sites') %>%
               ungroup() %>%
               gather(type, value, min, max),
             aes(yintercept = value),
             lty = 2,
             color = 'darkgray') +
  theme_minimal()

>>>>

All sites fell within the range of steelhead, but many of them were outside the range of spring/summer Chinook salmon. However, defining the range of sp/su Chinook is tricky, and it is unclear if the reported data for Chinook salmon includes only sites considered with the range of Chinook, or if we should exclude data from some sites based on our best understanding of the species' range. Currently, we are excluding sites outside the Chinook domain.

@@

# which of those sites are in Chinook domain?
chnk_sites = chnk_samps %>%
  as_Spatial() %>%
  maptools::snapPointsToLines(points = .,
                              lines = chnk_domain %>%
                                mutate(id = 1:n()) %>%
                                select(id, MPG) %>%
                                as_Spatial(),
                              maxDist = max_snap_dist,
                              withAttrs = TRUE,
                              idField = 'id') %>%
  as('sf') %>%
  select(-nearest_line_id, -snap_dist)

fish_hab %<>%
  filter(Site %in% chnk_sites$Site)

@@

```{r habitat-covariates}
sel_hab_mets = tibble(Species = c('Chinook'),
                      Metric = c("WetWDRat_Avg",
                                 "CU_Freq",
                                 "DpthWet_SD",
                                 "WetBraid",
                                 "FishCovTotal",
                                 "RipCovUstory",
                                 'WetWdth_Int',
                                 'Q',
                                 "SubLT6",
                                 "SubD16",
                                 "avg_aug_temp",
                                 "LWFreq_Wet"))
                      # 
                      # Metric = c('UcutArea_Pct',
                      #            'FishCovNone',
                      #            'SubEstGrvl',
                      #            'FstTurb_Freq',
                      #            'FstNT_Freq',
                      #            'CU_Freq',
                      #            'SlowWater_Pct',
                      #            'NatPrin1',
                      #            'DistPrin1',
                      #            'avg_aug_temp',
                      #            # 'Sin_CL',
                      #            'Sin',
                      #            'WetWdth_CV',
                      #            'WetBraid',
                      #            'WetSC_Pct',
                      #            'Q',
                      #            'WetWdth_Int',
                      #            'LWFreq_Wet',
                      #            'LWVol_WetFstTurb'))

@@

pred_hab_sites_chnk = pred_hab_sites %>%
  filter(!is.na(LON_DD)) %>%
  st_as_sf(coords = c('LON_DD', 'LAT_DD'),
           crs = 4326) %>%
  st_transform(crs = st_crs(chnk_domain)) %>%
  as_Spatial() %>%
  maptools::snapPointsToLines(chnk_domain %>%
                                mutate(id = 1:n()) %>%
                                select(id, MPG) %>%
                                as_Spatial(),
                              maxDist = snap_dist,
                              withAttrs = T,
                              idField = 'id') %>%
  as('sf') %>%
  as_tibble() %>%
  select(-nearest_line_id, -snap_dist, -geometry)

@@

# compare range of covariates from model dataset and prediction dataset
range_comp = bind_rows(gaa_all %>%
                         filter(!Site %in% unique(pred_hab_sites$Site)) %>%
                         select(Site,
                                one_of(gaa_num)) %>%
                         gather(Metric, value, -Site) %>%
                         mutate(Source = 'non-CHaMP Sites'),
                       gaa_all %>%
                         filter(Site %in% unique(pred_hab_sites$Site)) %>%
                         select(Site,
                                one_of(gaa_num)) %>%
                         distinct() %>%
                         gather(Metric, value, -Site) %>%
                         mutate(Source = 'CHaMP Sites')) %>%
  mutate_at(vars(Source, Metric),
            list(as.factor))

range_max = range_comp %>%
  group_by(Metric, Source) %>%
  summarise_at(vars(value),
               tibble::lst(min, max),
               na.rm = T) %>%
  filter(Source == 'CHaMP Sites') %>%
  ungroup() %>%
  gather(type, value, -Metric, -Source)

covar_range_p = range_comp %>%
  ggplot(aes(x = Source,
             y = value,
             fill = Source)) +
  geom_boxplot() +
  facet_wrap(~ Metric,
             scales = 'free') +
  geom_hline(data = range_comp %>%
               group_by(Metric, Source) %>%
               summarise_at(vars(value),
                            tibble::lst(min, max),
                            na.rm = T) %>%
               filter(Source == 'CHaMP Sites') %>%
               ungroup() %>%
               gather(type, value, min, max),
             aes(yintercept = value),
             lty = 2,
             color = 'darkgray') +
  theme_minimal()
